---
title: "Data overview and summary"
output:
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document highlights the data we will use for the hackathon. It describes the variables contained in the data and gives some preliminary plots
of the users time series

```{r load_data, echo = FALSE, warning=FALSE, message=FALSE, include=FALSE}
library(ggplot2)
library(reshape2)
getwd()
setwd('/Users/tnecamp/Box Sync/Intern-Health-Study-2014-15-cohort/Tim/All_2015_daily_data/2015 daily data/')
load("TN_merged_baseline_activity_sleep_mood_2015_ALL.Rdata")
full_2015_data$days_intern = full_2015_data$date - full_2015_data$StartDate_BS1
```

## Summary of data

For the purposes of the hackathon, we will be using the 2015 cohort data (the 2014 data has roughly the same variables but collected for half the year).

The data can be found here:

Intern-Health-Study-2014-15-cohort/Tim/All_2015_daily_data/2015 daily data/TN_merged_baseline_activity_sleep_mood_2015_ALL.Rdata

The data contains measurements on 23 interns in the intern health study. Data was collected on each intern roughly from 4/1/2015 to 7/1/2016. The
users started their internship July, roughly. The users and their start dates are given by:

```{r, echo=FALSE}
unique(full_2015_data[,c("userid", "StartDate_BS1")])
```


There are several baseline characterstics. Those are
```{r, echo=FALSE}
names(full_2015_data)[24:29]
```

There is also daily collected data. That data is collected in 3 categories.

Daily Mood- A self-reported score from 1-10

Daily Activity- which has variables
```{r, echo=FALSE}
names(full_2015_data)[4:20]
```

Daily Sleep- which has variables
```{r, echo=FALSE}
names(full_2015_data)[21:23]
```


Lastly, we conducted 6 survey on these users throughout the study. In the survey we asked the PHQ9 plus addtional information about 
sleep and work.
The question by question responses are given by
```{r, echo=FALSE}
names(full_2015_data)[30:135]
```

day_intern was also created to indicate the number of days they have been in the intership (negative values for before the internship).

From this point forward we will focus on exploring the daily measured data. Specifically we will look at daily mood, daily sleep total (TotalMinutesAsleep), and
daily activity total (TotalDistance in miles)

## Histograms

Here we give overall histograms across all users.

```{r hist, echo=FALSE}
hist(full_2015_data$mood, main = "Daily Mood")
hist(full_2015_data$TotalMinutesAsleep/60, main = "Daily hours of sleep")
hist(full_2015_data$TotalDistance, main = "Daily distance (miles)")
```



```{r correlation, echo=FALSE}
## get zero time lag correlations between different variables
all_users = unique(full_2015_data$userid)
cor_df = matrix(nrow = length(all_users), ncol = 3)
i=1
for(users_subset in all_users){
  data_subset = full_2015_data[(full_2015_data$userid %in% users_subset),]
  cor_df[i,1] = cor(data_subset$TotalMinutesAsleep, data_subset$mood, use = "complete.obs")
  cor_df[i,2] = cor(data_subset$TotalMinutesAsleep, data_subset$TotalDistance, use = "complete.obs")
  cor_df[i,3] = cor(data_subset$TotalDistance, data_subset$mood, use = "complete.obs")
  i=i+1
}
cor_df = data.frame(cor_df, row.names= all_users)
names(cor_df) = c("cor sleep_mood", "cor sleep_distance", "cor distance_mood")
cor_df$userid = all_users
cor_df_melt = melt(cor_df, id.vars = "userid")
ggplot(data = cor_df_melt, aes(x = userid, y = value)) + geom_point(aes(color = variable))
```


## Time series plots

Here we plot the times series for each user. The blue time series is always mood!

```{r mood_sleep, echo=FALSE, fig.asp = .5}
colormood <- "blue"
colorsleep <- "deeppink"
coloract <- "darkgreen"
labsize <- 1
ticksize <- 1
for(user_i in unique(full_2015_data$userid)){
  data_subset = full_2015_data[full_2015_data$userid ==user_i,]
  with(data_subset,plot(days_intern,mood,type="l",pch=16,col=colormood,xlab="Day in Study",
                        cex.lab=labsize, ylab="Mood self rating", cex.axis=ticksize,
                        main = paste("mood and sleep for user", user_i)))
  par(new=T)
  with(data_subset,plot(days_intern,TotalMinutesAsleep/60,type="l",pch=16,col=colorsleep,axes=F,xlab=NA,ylab=NA))
  axis(side=4,cex.axis=ticksize)
  mtext(side=4,line=3,"Hours Asleep",cex=labsize)
}
```
```{r mood_act, echo=FALSE, fig.asp = .5}
for(user_i in unique(full_2015_data$userid)){
  data_subset = full_2015_data[full_2015_data$userid ==user_i,]
  with(data_subset,plot(days_intern,mood,type="l",pch=16,col=colormood,xlab="Day in Study",
                        cex.lab=labsize, ylab="Mood self rating", cex.axis=ticksize,
                        main = paste("mood and activity for user", user_i)))
  par(new=T)
  with(data_subset,plot(days_intern,TotalDistance,type="l",pch=16,col=coloract,axes=F,xlab=NA,ylab=NA))
  axis(side=4,cex.axis=ticksize)
  mtext(side=4,line=3,"Total Distance (miles)",cex=labsize)
}
```

## Autocorrelation

Here we look at the autocorrelation of each time series for each user

```{r acf, echo=FALSE}
## Check that data is sorted
#for(user_i in unique(full_2015_data$userid)){
#  data_subset = full_2015_data[full_2015_data$userid ==user_i,]
#  print(sum(sort(data_subset[,2]) != data_subset[,2]))
#}
par(mfrow = c(2,3))
print("For mood")
for(user_i in unique(full_2015_data$userid)){
  data_subset = full_2015_data[full_2015_data$userid ==user_i,]
  cur_acf = acf(data_subset$mood, na.action = na.pass, plot = FALSE)
  plot(cur_acf, main = paste("mood autocorr", user_i), ylim = c(-.1, .3))
}
par(mfrow = c(2,3))
print("For Sleep")
for(user_i in unique(full_2015_data$userid)){
  data_subset = full_2015_data[full_2015_data$userid ==user_i,]
  cur_acf = acf(data_subset$TotalMinutesAsleep, na.action = na.pass, plot = FALSE)
  plot(cur_acf, main = paste("sleep autocorr", user_i), ylim = c(-.1, .3))
}
par(mfrow = c(2,3))
print("For Distance")
for(user_i in unique(full_2015_data$userid)){
  data_subset = full_2015_data[full_2015_data$userid ==user_i,]
  cur_acf = acf(data_subset$TotalDistance, na.action = na.pass, plot = FALSE)
  plot(cur_acf, main = paste("dist autocorr", user_i), ylim = c(-.1, .3))
}
```

## Crosscorrelation

Here we look at the crosscorrelation with mood and (sleep or distance) of each time series for each user. 

For sleep, mood: A lag of u is cor(Sleep[t+u], Mood[t])

For dist, mood: A lag of u is cor(Distance[t+u], Mood[t])

```{r ccf, echo=FALSE}
par(mfrow = c(2,3))
print("For Sleep and mood")
for(user_i in unique(full_2015_data$userid)){
  data_subset = full_2015_data[full_2015_data$userid ==user_i,]
  cur_ccf = ccf(data_subset$TotalMinutesAsleep, data_subset$mood, na.action = na.pass, plot = FALSE)
  plot(cur_ccf, main = paste("sleep, mood ccf", user_i))
}
par(mfrow = c(2,3))
print("For distance and mood")
for(user_i in unique(full_2015_data$userid)){
  data_subset = full_2015_data[full_2015_data$userid ==user_i,]
  cur_ccf = ccf(data_subset$TotalDistance, data_subset$mood, na.action = na.pass, plot = FALSE)
  plot(cur_ccf, main = paste("distance, mood ccf", user_i))
}
```






