---
title: "Vector Autoregression Results"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model Results from Fitting Vector Autoregressive models to every individual in the 2015 data

What is our model:

$Mood_t = \beta_0 + \beta_1 t + \beta_{2,1} Mood_{t-1} + \beta_{3,1} Sleep_{t} + \beta_{4,1} Activity_{t} +
... + \beta_{2,L} Mood_{t-L} + \beta_{3,L} Sleep_{t+1-L} + \beta_{4,L} Activity_{t+1-L}$

Here $Mood_t$ is the raw mood scores, $Sleep_t$ is that night's sleep in minutes, $Activity_t$ is logged activity for the day.
So I'm regressing the mood at 8 pm on the activity for that day, the sleep for the night before, the mood from the previous day
and I continue for L days into the past. L, the lag is chosen based on model fit.

Missing data- to deal with missing data we just do mean imputation. This is just a first pass, we could 
do something fancier later. Also, there are some users with lots of missingness, there results
may not be meaningful

```{r gettable, echo = FALSE, include = FALSE}
library(vars)
library(knitr)

setwd('/Users/tnecamp/Box Sync/Intern-Health-Study-2014-15-cohort/Tim/All_2015_daily_data/2015 daily data/')
load("TN_merged_baseline_activity_sleep_mood_2015_ALL.Rdata")


cur_data = full_2015_data[full_2015_data$userid == 110152, ]
missing_df = data.frame(matrix(nrow= 23, ncol = 4))
names(missing_df) = c('% mood NA', '% sleep NA', '% dist NA', 'time points')

model_summary_df = data.frame(matrix(nrow= 23, ncol = 3))
names(model_summary_df) = c('lags', 'R sq', 'adjusted R sq')

coeffs_df = data.frame(matrix(nrow= 23, ncol = 16))
coeffs_df_condensed =  data.frame(matrix(nrow= 23, ncol = 10))
names(coeffs_df_condensed) = c('mood1', 'p-val', 'sleep1', 'p-val', 'dist1', 'p-val',
                               'interc', 'p-val', 'trend', 'p-val')

intern_data = full_2015_data[full_2015_data$days_intern >= 0, ]

intern_data$mood_na = is.na(intern_data$mood)
mood_na_counts = aggregate(intern_data$mood_na, by = list(intern_data$userid), FUN = sum )
order_user = mood_na_counts$Group.1[order(mood_na_counts$x)]

for(user_i in 1:23){ #user_i indicates index of user
cur_user = order_user[user_i]
row.names(missing_df)[user_i] = cur_user
row.names(model_summary_df)[user_i] = cur_user
row.names(coeffs_df)[user_i] = cur_user
row.names(coeffs_df_condensed)[user_i] = cur_user

cur_data = full_2015_data[full_2015_data$userid == cur_user, ]
dim(cur_data)
xx = 0
for(i in 1:(nrow(cur_data)-1)){
  xx = xx + (as.numeric(cur_data$date[i] - cur_data$date[i+1]) == -1)
}
print(xx == (nrow(cur_data) - 1))
## so this being true indicates that they are all ordered and I have no date gaps. I think I can turn it into a time series now
x1 = ts(cur_data[cur_data$days_intern >= 0, 'mood'])
x2 = ts(cur_data[cur_data$days_intern >= 0, 'TotalMinutesAsleep'], start = 0)
x_new = ts.intersect(x1,x2)
cur_data$sqrtDist = sqrt(cur_data$TotalDistance)
x3 = ts(cur_data[cur_data$days_intern >= 0, 'sqrtDist'], start = 0)
cur_ts = ts.intersect(x_new, x3)
colnames(cur_ts) = c('mood', 'TotalMinutesAsleep', 'sqrtDist')

missing_df[user_i, '% mood NA'] = sum(is.na(cur_ts[,'mood']))/(nrow(cur_ts)) 
missing_df[user_i, '% sleep NA'] = sum(is.na(cur_ts[,'TotalMinutesAsleep']))/(nrow(cur_ts)) 
missing_df[user_i, '% dist NA'] = sum(is.na(cur_ts[,'sqrtDist']))/(nrow(cur_ts))
missing_df[user_i, 'time points'] = nrow(cur_ts)

# mean imputation
cur_ts[is.na(cur_ts[,'mood']), 'mood'] = mean(cur_ts[,'mood'], na.rm = TRUE)
cur_ts[is.na(cur_ts[,'sqrtDist']), 'sqrtDist'] = mean(cur_ts[,'sqrtDist'], na.rm = TRUE)
cur_ts[is.na(cur_ts[,'TotalMinutesAsleep']), 'TotalMinutesAsleep'] = mean(cur_ts[,'TotalMinutesAsleep'], na.rm = TRUE)

cur_fit = VAR(cur_ts[,c('mood', 'TotalMinutesAsleep', 'sqrtDist')], type = 'both', lag.max = 5) ## you don't need to have both days intern and trend, one or the other

cur_sum = summary(cur_fit$varresult$mood)
cur_coeffs = as.vector(t(cur_sum$coefficients)[c(1,4), ]) # #this gives  the coefficients I want. there is a slight problem if my lag is bigger than one, then there is 
## no alignment when putting it all together in a matrix

cur_lag = cur_fit$p
temp = 6*cur_lag + 1
coeffs_df_condensed[user_i, ] = cur_coeffs[c(1:6, temp:(temp+3))]

length(cur_coeffs) = 16
coeffs_df[user_i, ] = cur_coeffs

model_summary_df[user_i, 'R sq'] = cur_sum$r.squared  # gets r squared value
model_summary_df[user_i, 'adjusted R sq'] = cur_sum$adj.r.squared # gets adjusted r squared value
model_summary_df[user_i, 'lags'] = cur_fit$p # this gets the lag
}

```

## Results of model fits for each user.

Each row represents a new users' fit.

Here we include the missingness in the data for each covariate, the number of data selected lags, the goodness of the fit (Rsq) and the 
coefficients for each lag and their corresponding p-values. The coefficients are just for the first lag to make the table reasonable sized, but there may have been more based
on the chosen number of lags.
The intercept is $\beta_0$ and the trend is $\beta_1$

```{r pressure, echo=FALSE}
kable(cbind(round(missing_df,2), round(model_summary_df,3), round(coeffs_df_condensed,3)))
## maybe use kable here
```

## Plots of time series and residuals

Here we plot the time series for mood, activity and sleep. We also have residual plots for each user. The plots include the fit, residuals over time, the acf and partial acf of the residuals

```{r user plots, echo=FALSE, fig.height = 8, fig.width = 10}
for(user_i in 1:23){ #user_i indicates index of user
  cur_user = order_user[user_i]
  print('user')
  print(cur_user)
  row.names(missing_df)[user_i] = cur_user
  row.names(model_summary_df)[user_i] = cur_user
  row.names(coeffs_df)[user_i] = cur_user
  row.names(coeffs_df_condensed)[user_i] = cur_user
  
  cur_data = full_2015_data[full_2015_data$userid == cur_user, ]
  ## so this being true indicates that they are all ordered and I have no date gaps. I think I can turn it into a time series now
  x1 = ts(cur_data[cur_data$days_intern >= 0, 'mood'])
  x2 = ts(cur_data[cur_data$days_intern >= 0, 'TotalMinutesAsleep'], start = 0)
  x_new = ts.intersect(x1,x2)
  cur_data$sqrtDist = sqrt(cur_data$TotalDistance)
  x3 = ts(cur_data[cur_data$days_intern >= 0, 'sqrtDist'], start = 0)
  cur_ts = ts.intersect(x_new, x3)
  colnames(cur_ts) = c('mood', 'TotalMinutesAsleep', 'sqrtDist')
  
  plot(cur_ts)
  
  # mean imputation
  cur_ts[is.na(cur_ts[,'mood']), 'mood'] = mean(cur_ts[,'mood'], na.rm = TRUE)
  cur_ts[is.na(cur_ts[,'sqrtDist']), 'sqrtDist'] = mean(cur_ts[,'sqrtDist'], na.rm = TRUE)
  cur_ts[is.na(cur_ts[,'TotalMinutesAsleep']), 'TotalMinutesAsleep'] = mean(cur_ts[,'TotalMinutesAsleep'], na.rm = TRUE)
  
  cur_fit = VAR(cur_ts[,c('mood', 'TotalMinutesAsleep', 'sqrtDist')], type = 'both', lag.max = 5) ## you don't need to have both days intern and trend, one or the other
  plot(cur_fit, names = 'mood', oma = c(0,0,0,0), mar= c(0,0,0,0))
  }
```
